---
title: "Spatio-temporal study distribution"
author: "Julianna Renzi"
date: "4/15/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: lumen
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load required packages
require(tidyverse) # data manipulation
require(here) # for relative paths
require(RColorBrewer) # for colors
require(patchwork) # combining figures
require(sf) # working with georeferenced objects
require(rnaturalearth) # map of the world
require(ggsn) # scale bar
```

# Make a plot for the number of studies over time

First bring in the correct dataset that has a unique line per study and a column noting the type of experiment (observational, experimental, or both)

```{r}
# bring in the data
time <- read_csv(here("data", "Corallivore_year_exptype.csv")) # this is one line for each study
```

Summarize the number of studies per year

```{r}
# get the count of publications per year
time_sum <- time %>% 
  group_by(Year, Exp_type) %>% 
  summarize(Publications = n()) 

# get the duration of the graph we want (from the first study to the last)
min_yr <- 1976
max_yr <- 2020

# create a dataframe of all possible values to join so that every year is accounted for (year=0 if there are no publications)
all_years <- data.frame(Year = c(seq(min_yr, max_yr, by = 1),
                                seq(min_yr, max_yr, by = 1),
                                seq(min_yr, max_yr, by = 1)),
                        Exp_type = c(rep("Experimental", times = max_yr-min_yr+1),
                                     rep("Observational", times = max_yr-min_yr+1),
                                     rep("Both", times = max_yr-min_yr+1))) 

# join the dataframes so there's a line for every year
time_full <- all_years %>% 
  full_join(time_sum, by = c("Year", "Exp_type")) %>% # join the two dataframes so there's a line for every year
  mutate(Publications = replace_na(Publications, 0)) # replace the NA's for 0 in the years where nothing was published

```

Then make the plot

```{r}
time_plt <- time_full %>% 
  ggplot(aes(x = Year, y = Publications, fill = fct_reorder(Exp_type, desc(Exp_type)))) + # fct_reorder() so both is the last in the legend (makes the most sense)
  geom_bar(stat = "identity") + # identity means to use the values in the cells for plotting
  ylab("Number of studies") +
  geom_hline(yintercept = 0) + # create an x-axis that is distinct
  theme_minimal()  +
  scale_fill_manual(name = "", # no legend title 
                    values = rev(brewer.pal(n = 3, name = "YlOrBr"))) +
  scale_x_continuous(breaks = seq(min_yr, max_yr, by = 2)) + # get higher resolution ticks on the x-axis
  theme(text = element_text(size = 15), legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1), # rotate axis labels
        axis.title.x = element_text(vjust = -1), # add padding fo titles (also need to adjust the margins)
        axis.title.y = element_text(vjust = 3),
        plot.margin = unit(c(5.5, 5.5, 10, 10), "points")) 

```

# Make a map of where studies were conducted

First bring in polygons of the world from `rnaturalearth` using `ne_countries` and polygons of marine ecoregions from Spalding et al. 2007 (downloaded from: https://geospatial.tnc.org/datasets/ed2be4cf8b7a451f84fd093c2e7660e3_0)

```{r}
# bring in world shape
world <- ne_countries(scale = "medium", returnclass = "sf")

# bring in shapefiles from Spalding et al. on marine bioregions
meow_ecos <- read_sf(here("data", "MEOW-TNC"), layer = "meow_ecos")

# check that projections match
st_crs(world) == st_crs(meow_ecos) # WGS 84
```

Then bring in the csv that has a line for each ecoregion in each study with the associated publication name

```{r}
# bring in the csv
ecoregions <- read_csv(here("data", "Corallivore_spatial.csv")) # there could be more than one ecoregion/study (multiple rows in the file) if it was conducted in more than one ecoregion

# get count of number of studies per ecoregion
ecoreg_sum <- ecoregions %>% 
  group_by(Ecoregion) %>% 
  summarize(Num_studies = n()) %>% 
  mutate(ECOREGION = Ecoregion) # make name match that in meow_ecos

# join to the Spalding et al. dataset for plotting
meow_pub <- meow_ecos %>% 
  left_join(ecoreg_sum, by = "ECOREGION")

```

Then plot the results

```{r}
ecoreg_plt <- ggplot() +
  # color the ecoregions based on the number of studies they have
  geom_sf(data = meow_pub, aes(fill = Num_studies), color = NA, size = 0) +
  geom_sf(data = world, fill = "black", color = "black", size = 0.05) + # background color for continents (black)
  theme_bw() +
  # format legend and color scheme using RColorBrewer
  scale_fill_gradient(name = "Number of studies", 
                      low = brewer.pal(n = 3, name = "YlOrRd")[1], 
                      high = brewer.pal(n = 3, name = "YlOrRd")[3], 
                      na.value = NA, breaks = seq(1, 14, by = 2)) +
  theme(text = element_text(size = 15)) + # increase text size
  coord_sf(xlim = c(-180, 180), ylim = c(-45, 45), expand = FALSE) + # crop the view so get a better look
  theme(panel.grid.major = element_line(colour = 'grey50', size = 0.2, linetype = 3)) + # add grid lines (no scale bar because projection would make one scale bar meaningless over the entire map)
  north(data = world, location = "bottomleft", scale = 0.1, symbol = 14, anchor = c(x = -180, y = -40)) # add a north arrow

```

# Combine figures to create Figure 1

Use `patchwork` to combine panels

```{r}
# to save as pdf() run this chunk without hashtags
# pdf("figures/Figure_1.pdf", width = 15, height = 5)
time_plt + ecoreg_plt +
   plot_annotation(tag_levels = list(c("(a)", "(b)")))

# dev.off()
```

# Get stats on spatial distribution

We quote some statistics of where studies were conducted. This code is how we got those

1. For ecoregions:

```{r}
# get the total number of unique papers
total.pubs <- time %>% 
  select(Publication) %>% 
  unique() %>% 
  count() %>% 
  as.numeric()

# get the total number of geographic observations (this is larger than the number of unique papers because some papers were conducted in multiple ecoregions)
val <- ecoreg_sum %>% 
  summarize(Total_studies = sum(Num_studies)) %>% 
  as.numeric()

# get the number of ecoregions that had studies
ecoreg_num <- ecoreg_sum %>% 
  summarize(Total_regions = n()) %>% 
  as.numeric()

# get the percentage of studies per region
perc_df <- ecoreg_sum %>% 
  group_by(ECOREGION) %>% 
  summarize(Percent = Num_studies/val*100,
            Num_std = Num_studies) %>% 
  arrange(-Percent)
```

2. By realm:

```{r}
# get count per realm
realm_pub <- meow_pub %>% 
  select(REALM, Num_studies) %>% 
  group_by(REALM) %>% 
  summarize(Num_studies = sum(Num_studies, na.rm = TRUE)) 

realm_nums <- realm_pub %>% 
  filter(Num_studies != 0) %>% 
  select(REALM, Num_studies) 

# get total number of regions
val_r <- realm_nums %>% 
  summarize(Total_studies = sum(Num_studies)) %>% 
  as.numeric()

realm_perc <- realm_nums %>% 
  group_by(REALM) %>% 
  summarize(Percent = Num_studies/val_r[1]*100,
            Num_std = Num_studies)

```



