---
title: "Corallivore-disease relationships"
author: "Julianna Renzi"
date: "4/15/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
require(tidyverse) # data manipulation
require(here) # for relative paths
require(RColorBrewer) # for colors
require(patchwork) # combining figures
```

# Correlations between disease and corallivores

Bring in the data that has a unique line for each corallivore-study combination

```{r}
corallivores <- read_csv(here("data", "Taxa_effects.csv")) %>% 
  select(-starts_with(match = "X")) # get rid of extra columns read_csv can bring in
```

Look at which families are correlated with disease

```{r}
corr_effect <- corallivores %>% 
  filter(!is.na(Correlation)) %>% 
  filter(Correlation != "None") %>% # only look at the ones with a correlation
  filter(Disease_general != "Microbiome") %>% # don't want to use this because it's too general
  group_by(Corallivore_family, Disease_general) %>% 
  summarize(N_pos = sum(Correlation == "Positive"),
            N_neg = sum(Correlation == "Negative")) 

```

Set a color scheme to use throughout

```{r}
asp_col <- brewer.pal(n = 9, name = "Spectral")[1]
bbd_col <- brewer.pal(n = 9, name = "Spectral")[2]
blc_col <- brewer.pal(n = 9, name = "Spectral")[3]
brb_col <- brewer.pal(n = 9, name = "Spectral")[4]
cya_col <- brewer.pal(n = 9, name = "Spectral")[5]
ga_col <- brewer.pal(n = 9, name = "Spectral")[6]
gen_col <- brewer.pal(n = 9, name = "Spectral")[7]
seb_col <- brewer.pal(n = 9, name = "Spectral")[8]
trem_col <- brewer.pal(n = 9, name = "Spectral")[9]
whi_col <- brewer.pal(n = 11, name = "Spectral")[11]

```


## Plot corallivore families that are positively correlated with specific diseases

```{r}
# positive
corr_pos_plt <- corr_effect %>% 
  filter(N_pos != 0) %>% # only studies with a positive relationship found
  ggplot(aes(x = ordered(Corallivore_family, levels = c("Muricidae", "Pomacentridae", "Amphinomidae", "Chaetodontidae", "Ovulidae", "Xanthidae", "Cryptochiridae", "Scaridae")), 
             y = N_pos, 
             fill = Disease_general)) +
    geom_bar(stat = "identity") +
    theme_classic() +
    scale_fill_manual(name = "Disease",
                        values = c("Aspergillosis" = asp_col, "BBD" = bbd_col, "Bleaching" = blc_col, "Cyanobacterial" = cya_col, "GA" = ga_col, "General" = gen_col, "SEB" = seb_col, "Trematodiasis" = trem_col, "White" = whi_col, "BrB" = brb_col)) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "none")  +
  ggtitle("(a) Positive correlations") +
  xlab("") +
  ylab("Number of observations") +
  scale_y_continuous(breaks = seq(0, 10, by = 2), limits = c(0, 10))
```


## Plot corallivore families that are negatively correlated with specific diseases

```{r}
# negative
corr_neg_plt <- corr_effect %>% 
  filter(N_neg != 0) %>% 
  ggplot(aes(x = fct_rev(Corallivore_family), 
             y = N_neg, 
             fill = Disease_general)) +
    geom_bar(stat = "identity") +
    theme_classic() +
    scale_fill_manual(name = "Disease",
                        values = c("Aspergillosis" = asp_col, "BBD" = bbd_col, "Bleaching" = blc_col, "Cyanobacterial" = cya_col, "GA" = ga_col, "General" = gen_col, "SEB" = seb_col, "Trematodiasis" = trem_col, "White" = whi_col, "BrB" = brb_col)) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "none")  +
  ggtitle("(b) Negative correlations") +
  xlab("") +
  ylab("") +
  scale_y_continuous(breaks = seq(0, 10, by = 2), limits = c(0, 10))
```

## Experimental studies showing corallivore-disease relationship

Plot studies that tested whether corallivores increase/decrease/have no effect on disease. Start by summarizing the data

```{r}
manip_effect <- corallivores %>% 
  filter(!is.na(Experimental_effect)) %>% # only studies that involved an experiment
  filter(Disease_general != "Microbiome") %>% # don't want to use this because it's too general
  group_by(Corallivore_family, Disease_general) %>% 
  summarize(N_inc = sum(Experimental_effect == "Increased"),
            N_dec = sum(Experimental_effect == "Decreased"),
            N_noeff = sum(Experimental_effect == "No effect")) 
```

Plot by increased/decreased. First by increased:

```{r}
# increased
manip_pos_plt <- manip_effect %>% 
  filter(N_inc != 0) %>% 
  ggplot(aes(x = ordered(Corallivore_family, 
                         levels = c("Muricidae", "Amphinomidae", "Acanthasteridae", "Cryptochiridae", "Trinchesiidae", "Chaetodontidae")), 
             y = N_inc, 
             fill = Disease_general)) +
    geom_bar(stat = "identity") +
    theme_classic() +
    scale_fill_manual(name = "Disease",
                        values = c("Aspergillosis" = asp_col, "BBD" = bbd_col, "Bleaching" = blc_col, "Cyanobacterial" = cya_col, "GA" = ga_col, "General" = gen_col, "SEB" = seb_col, "Trematodiasis" = trem_col, "White" = whi_col, "BrB" = brb_col)) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "none")  +
  ggtitle("(c) Increased disease") +
  xlab("") +
  ylab("Number of observations") +
  scale_y_continuous(breaks = seq(0, 10, by = 2), limits = c(0, 10))

```

Then for decreased:

```{r}
# decreased
manip_neg_plt <- manip_effect %>% 
  filter(N_dec != 0) %>% 
  ggplot(aes(x = Corallivore_family, 
             y = N_dec, 
             fill = Disease_general)) +
    geom_bar(stat = "identity") +
    theme_classic() +
    scale_fill_manual(name = "Disease",
                        values = c("Aspergillosis" = asp_col, "BBD" = bbd_col, "Bleaching" = blc_col, "Cyanobacterial" = cya_col, "GA" = ga_col, "General" = gen_col, "SEB" = seb_col, "Trematodiasis" = trem_col, "White" = whi_col, "BrB" = brb_col)) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "none")  +
  ggtitle("(d) Decreased disease") +
  xlab("") +
  ylab("") +
  scale_y_continuous(breaks = seq(0, 10, by = 2), limits = c(0, 10))
```

Then for no effect:

```{r}
# no effect
manip_noEff_plt <- manip_effect %>% 
  filter(N_noeff != 0) %>% 
  ggplot(aes(x = ordered(Corallivore_family,
                         levels = c("Chaetodontidae", "Muricidae", "Amphinomidae", "Pomacentridae")), 
             y = N_noeff, 
             fill = Disease_general)) +
    geom_bar(stat = "identity") +
    theme_classic() +
    scale_fill_manual(name = "Disease",
                        values = c("Aspergillosis" = asp_col, "BBD" = bbd_col, "Bleaching" = blc_col, "Cyanobacterial" = cya_col, "GA" = ga_col, "General" = gen_col, "SEB" = seb_col, "Trematodiasis" = trem_col, "White" = whi_col, "BrB" = brb_col)) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "none")  +
  ggtitle("(e) No effect") +
  xlab("") +
  ylab("") +
  scale_y_continuous(breaks = seq(0, 10, by = 2), limits = c(0, 10))
```

# get one legend for them all

```{r}
# plot something with all of the colors (this is very sloppily done)
legend_plt <- corr_effect %>% 
  as.data.frame() %>% 
  add_row("Corallivore_family" = "Test", "Disease_general" = "BrB", "N_pos" = 8, "N_neg" = 8) %>% 
  ggplot(aes(x = Corallivore_family, 
             y = N_pos, 
             fill = Disease_general)) +
    geom_bar(stat = "identity") +
    theme_classic() +
    scale_fill_manual(name = "Disease",
                        values = c("Aspergillosis" = asp_col, "BBD" = bbd_col, "Bleaching" = blc_col, "Cyanobacterial" = cya_col, "GA" = ga_col, "General" = gen_col, "SEB" = seb_col, "Trematodiasis" = trem_col, "White" = whi_col, "BrB" = brb_col)) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1))  +
  xlab("") +
  ylab("Number of observations") +
  scale_y_continuous(breaks = seq(0, 10, by = 2), limits = c(0, 10))

# extract legend
# used code from: https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend <- g_legend(legend_plt)
```

# Combine figures

```{r}
# pdf(here("figures/Figure_2.pdf"), height = 7.5, width = 11)

(corr_pos_plt + corr_neg_plt + mylegend) /
(manip_pos_plt + manip_neg_plt + manip_noEff_plt) & theme(text = element_text(size = 18), plot.title = element_text(size = 18))        


# dev.off()
```



